Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-11-20 paper.qmd
Added results for different UMAP parameter choices

@@

nldr1c <- umap_pbmc1 |>
  ggplot(aes(x = UMAP1,
             y = UMAP2)) +
  geom_point(alpha=0.1, size=1, colour='#a65628') +
  interior_annotation("a")

@@

nldr2c <- umap_pbmc2 |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1, colour='#999999') +
  interior_annotation("b")

@@

nldr3c <- umap_pbmc3 |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1, colour='#f781bf') +
  interior_annotation("c")
```

@@

wrap_plots(nldr1c, nldr2c, nldr3c, ncol = 3)
```

>>>>

result_umap <- result_umap |>
  mutate(plot = c("b", "i", "h", "g", "f", "a", "e", "c", "d")) |>
  arrange(number_dubious_cells)

@@

nldr1c <- umap_pbmc1 |>
  ggplot(aes(x = UMAP1,
             y = UMAP2)) +
  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
  interior_annotation("a")

@@

nldr2c <- umap_pbmc2 |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1, colour='#a65628') +
  interior_annotation("b")

@@

nldr3c <- umap_pbmc3 |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
  interior_annotation("c")

umap_pbmc4 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_30_min_dist_0.99.rds")
umap_pbmc4 <- as_tibble(umap_pbmc4)
names(umap_pbmc4) <- c("UMAP1", "UMAP2")

nldr4c <- umap_pbmc4 |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1, colour='#8dd3c7') +
  interior_annotation("d")

umap_pbmc5 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_5_min_dist_0.99.rds")
umap_pbmc5 <- as_tibble(umap_pbmc5)
names(umap_pbmc5) <- c("UMAP1", "UMAP2")

nldr5c <- umap_pbmc5 |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1, colour='#999999') +
  interior_annotation("e")

umap_pbmc6 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_15_min_dist_0.3.rds")
umap_pbmc6 <- as_tibble(umap_pbmc6)
names(umap_pbmc6) <- c("UMAP1", "UMAP2")

nldr6c <- umap_pbmc6 |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1, colour='#377eb8') +
  interior_annotation("f")

umap_pbmc7 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_5_min_dist_0.3.rds")
umap_pbmc7 <- as_tibble(umap_pbmc7)
names(umap_pbmc7) <- c("UMAP1", "UMAP2")

nldr7c <- umap_pbmc7 |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1, colour='#f781bf') +
  interior_annotation("g")

umap_pbmc8 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_30_min_dist_0.01.rds")
umap_pbmc8 <- as_tibble(umap_pbmc8)
names(umap_pbmc8) <- c("UMAP1", "UMAP2")

nldr8c <- umap_pbmc8 |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1, colour='#984ea3') +
  interior_annotation("h")

umap_pbmc9 <- read_rds("data/pbmc3k/pbmc_scdeed_umap_n_neighbors_15_min_dist_0.01.rds")
umap_pbmc9 <- as_tibble(umap_pbmc9)
names(umap_pbmc9) <- c("UMAP1", "UMAP2")

nldr9c <- umap_pbmc9 |>
  ggplot(aes(x = UMAP1,
             y = UMAP2))+
  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
  interior_annotation("i")
```

```{r}
#| label: combine-data-pbmc-umap

error_pbmc_umap <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_30_min_dist_0.3.rds")
error_pbmc_umap2 <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_5_min_dist_0.01.rds")
error_pbmc_umap3 <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_15_min_dist_0.99.rds")
error_pbmc_umap4 <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_30_min_dist_0.99.rds")
error_pbmc_umap5 <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_5_min_dist_0.99.rds")
error_pbmc_umap6 <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_15_min_dist_0.3.rds")
error_pbmc_umap7 <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_5_min_dist_0.3.rds")
error_pbmc_umap8 <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_30_min_dist_0.01.rds")
error_pbmc_umap9 <- read_rds("data/pbmc3k/error_scdeed_pbmc_umap_15_min_dist_0.01.rds")

error_pbmc <- bind_rows(error_pbmc_umap, 
                        error_pbmc_umap2,
                        error_pbmc_umap3,
                        error_pbmc_umap4,
                        error_pbmc_umap5,
                        error_pbmc_umap6,
                        error_pbmc_umap7,
                        error_pbmc_umap8,
                        error_pbmc_umap9)

```{r}
#| label: error-comp-pbmc-umap

error_umap_plot_pbmc <- ggplot(error_pbmc, 
                          aes(x = a1, 
                              y = MSE, 
                              colour = method)) + 
  geom_point(size = 0.8) +
  geom_line(linewidth = 0.3) + 
  # geom_vline(xintercept = 15, linetype="solid", 
  #            color = "black", linewidth=0.8, alpha = 0.5) +
  scale_x_continuous(breaks = sort(unique(error_pbmc$a1))[c(1, 5, 9, 13, 17, 21, 26)]) +
  scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#8dd3c7','#a65628','#f781bf','#999999')) +
  scale_y_log10() +
  ylab("log(MSE)") +
  xlab(expression(paste("binwidth (", a[1], ")"))) +
  theme_minimal() +
  theme(panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line(),
        legend.position = "none",
        axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 7),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7))

@@

free(error_umap_plot_pbmc) + wrap_plots(nldr1c, nldr2c, nldr3c, nldr4c, nldr5c, nldr6c, nldr7c, nldr8c, nldr9c, ncol = 2)
```
